---
title: "movie.dataunderstanding2"
author: "Jingnan Wang"
date: "2021/10/5"
output: html_document
---

```{r}
library(ggplot2)
movie = read.csv("cleaned_data_v2.csv")
movie
```
To analyze the potential effect of multicolinearity in the dataset, we choose to construct a correlation matrix including all the numerical variables in the dataset (all variables except content rating and color) and then visualize with a correlation plot.

```{r}
##correlation matrix
use<-movie[ -c(4:5) ]
cor_matrix = cor(use, method = "pearson", use = "complete.obs")

library(corrplot)
```

Since we have 39 variables which is hard to visualize in a single plot, we chose to only select variable pairs whose correlations are over 0.5. Then we construct a correlation plot.

```{r}
##finding correlations over 0.5 and plot the correlation matrix
refinedcorr <- function(data=use,sig=0.5){
  df_cor<-cor(use)
  corr <- cor(df_cor)
  corr[lower.tri(corr,diag=TRUE)] <- NA 
  corr[corr == 1] <- NA 
  corr <- as.data.frame(as.table(corr))
  corr <- subset(corr, abs(Freq) > sig) 
  corr <- corr[order(-abs(corr$Freq)),] 
  print(corr)
  corr1 <- reshape2::acast(corr, Var1~Var2, value.var="Freq")
  corrplot(corr1, is.corr=FALSE, tl.col="black", na.label=" ")
}
refinedcorr()
```

From the result, we see some multicolinearity issues in our dataset which may further influence our modeling. For example, actor_1_facebook_likes and cast_facebook_likes, Family and Animation. Some high correlations indeed seem reasonable, and some do not. 

Then we consider to look at the distribution for the profit in millions of dollars for our movies. We construct a historgram, and find the distribution to be right-skewed, suggesting certain movies have extremely high profit than other movies and act like outliers.

```{r}
##plot the histogram
hist(movie$Profit/1000000,main="Histogram for Profit", 
     xlab="million in dollars")
mean(movie$Profit)
median(movie$Profit)
```

When we choose the profit for a movie as the business metric to analyze, we use budget of the movie and its gross. We really want to find whether a highly budgeted movie will ensure a high return in profit. To analyze this question, we fit a scatter plot for the profit of the movie against its budget. 


```{r}
## scatter plot for the relationship between budget and profit
budget_in_millions<-movie$budget_corrected/1000000
profit_in_millions<-movie$Profit/1000000
ggplot(movie, aes(budget_in_millions,profit_in_millions))+
  geom_point(position = "jitter",alpha = 0.3,color = "blue",shape = ".")+
  geom_smooth(color ="black")
```

From the scatter plot, we see a general and increasing trend that a movie's profit is positively related to its budget, suggesting that higher the cost of the movie will generate a higher profit.

We are also interested to see the relationship between a movie's genre and its profit. Which movie genres will be likely to generate higher profit?
To analyze this point, we calculated the average profit for movies in each genre and aim to compare the average profit in a barplot.


```{r}
##calculate the average profit for each genre's movies and fit a barplot of profit by genre

fantasy<-sum(movie[movie$Fantasy>=1,"Profit"])/length(movie[movie$Fantasy>=1,"Fantasy"])

horror<-sum(movie[movie$Horror>=1,"Profit"])/length(movie[movie$Horror>=1,"Horror"])
crime<-sum(movie[movie$Crime>=1,"Profit"])/length(movie[movie$Crime>=1,"Crime"])
history<-sum(movie[movie$History>=1,"Profit"])/length(movie[movie$History>=1,"History"])
biography<-sum(movie[movie$Biography>=1,"Profit"])/length(movie[movie$Biography>=1,"Biography"])
family<-sum(movie[movie$Family>=1,"Profit"])/length(movie[movie$Family>=1,"Family"])
adventure<-sum(movie[movie$Adventure>=1,"Profit"])/length(movie[movie$Adventure>=1,"Adventure"])
animation<-sum(movie[movie$Animation>=1,"Profit"])/length(movie[movie$Animation>=1,"Animation"])
sport<-sum(movie[movie$Sport>=1,"Profit"])/length(movie[movie$Sport>=1,"Sport"])
war<-sum(movie[movie$War>=1,"Profit"])/length(movie[movie$War>=1,"War"])
musical<-sum(movie[movie$Musical>=1,"Profit"])/length(movie[movie$Musical>=1,"Musical"])
drama<-sum(movie[movie$Drama>=1,"Profit"])/length(movie[movie$Drama>=1,"Drama"])
music<-sum(movie[movie$Music>=1,"Profit"])/length(movie[movie$Music>=1,"Music"])
action<-sum(movie[movie$Action>=1,"Profit"])/length(movie[movie$Action>=1,"Action"])
romance<-sum(movie[movie$Romance>=1,"Profit"])/length(movie[movie$Romance>=1,"Romance"])
western<-sum(movie[movie$Western>=1,"Profit"])/length(movie[movie$Western>=1,"Western"])
documentary<-sum(movie[movie$Documentary>=1,"Profit"])/length(movie[movie$Documentary>=1,"Documentary"])
mystery<-sum(movie[movie$Mystery>=1,"Profit"])/length(movie[movie$Mystery>=1,"Mystery"])
sci_fi<-sum(movie[movie$Sci.Fi>=1,"Profit"])/length(movie[movie$Sci.Fi>=1,"Sci.Fi"])
thriller<-sum(movie[movie$Thriller>=1,"Profit"])/length(movie[movie$Thriller>=1,"Thriller"])
comedy<-sum(movie[movie$Comedy>=1,"Profit"])/length(movie[movie$Comedy>=1,"Comedy"])
genre_others<-sum(movie[movie$Genre_Others>=1,"Profit"])/length(movie[movie$Genre_Others>=1,"Genre_Others"])

averageprofit<-c(horror,fantasy,crime,history,biography,family,adventure,animation,sport,war,musical,drama,music,action,romance,western,documentary,mystery,sci_fi,thriller,comedy,genre_others)

genre<-c("Fantasy","Horror","Crime","History","Biography","Family","Adventure","Animation","Sport","War","Musical","Drama","Music","Action","Romance","Western","Documentary","Mystery","Sci_fi","Thriller","Comedy","Genre_others")

barplot(averageprofit,names.arg=genre,xlab="Genre",ylab="Profit",col="blue",width=0.05,cex.names=0.45,
main="Bar plot for profit by genre",border="red")

```

From the barplot, we see that animation,adventure movies are among the genres which generated the highest profit, while documentary and sport movies generally generate lower profit.
